---
title: "Times Higher Education World University Rankings"
author: "Colin Dempsey"
format: html
editor: visual
---

```{r} 

library(tidyverse)
library(reactable)
library(crosstalk)
library(plotly)
```

# Problem 1A

Read in data.

```{r}

times <- read_csv(
  "https://manctsui.github.io/datasets/timesData.csv",
  show_col_types = FALSE
) |>
  mutate(
    rank = readr::parse_number(world_rank)
  ) |>
  group_by(country) |>
  mutate(
    Number_of_Universities = n()
  ) |>
  ungroup()

head(times)
names(times)
```

Find top three countries with the most universities.

```{r }
country_counts <- times |>
  distinct(country, Number_of_Universities) |>
  arrange(desc(Number_of_Universities)) |>
  slice_head(n = 3)
```

Display as a reactable.

```{r }
reactable(
  country_counts,
  searchable = FALSE,
  pagination = FALSE,
  bordered = TRUE,
  striped = TRUE,
  highlight = TRUE,
  columns = list(
    country = colDef(
      name = "Country"
    )
  )
)
```

# Problem 1B

Display the results from Problem 1A as a bar chart.

```{r }
ggplot(
  country_counts,
  aes(
    x = reorder(country, Number_of_Universities),
    y = Number_of_Universities
  )
) +
  geom_col() +
  coord_flip() +
  labs(
    title = "Three Countries with the Most Universities",
    x = "Country",
    y = "Number of Universities"
  ) +
  theme_minimal()
```

# Problem 2

Display top 50 universities in 2011.

```{r }
top_50_2011 <- times |>
  filter(year == 2011) |>
  arrange(rank) |>
  slice_head(n = 50) |>
  select(
    Rank = rank,
    University_Name = university_name
  )
```

```{r }
reactable(
  top_50_2011,
  columns = list(
    University_Name = colDef(
      name = "University Name"
    )
  )
)
```

# Problem 2B

The rank column by itself would not be a good candidate for a SharedData key. The data set contains multiple years and thus the same rank appears in multiple years so rank alone cannot uniquely identify an observation. Rank and year is a better choice.

add key column using mutate
```{r}
times <- times |>
  mutate(
    ranking_key = paste(year, rank, sep = "-")
  )
```

# Problem 2C

Create SharedData object.

```{r }
shared_times <- SharedData$new(
  times,
  key = ~ranking_key,
  group = "university_rankings"
)
```

Create a filter slider based on university rank.

```{r }
filter_slider(
  id = "rank_filter",
  label = "University Rank:",
  sharedData = shared_times,
  column = ~rank,
  step = 1
)
```

# Problem 3

Restrict dataset 2016.

```{r }
times_2016 <- times |>
  filter(year == 2016)
```

```{r }
shared_2016 <- SharedData$new(
  times_2016,
  key = ~ranking_key,
  group = "rankings_2016"
)
```

## Number of Students vs. Student-Staff Ratio

```{r }
plot_ly(
  data = shared_2016,
  x = ~num_students,
  y = ~student_staff_ratio,
  type = "scatter",
  mode = "markers"
)
```

## Student-Staff Ratio vs. Teaching Score

```{r }
plot_ly(
  data = shared_2016,
  x = ~student_staff_ratio,
  y = ~teaching,
  type = "scatter",
  mode = "markers"
)
```

