---

title: "Assignment on Theoretical Data Transformations"
author: "Colin Dempsey"
date: "r Sys.Date()"
output: html_document
---

# Problem 1

```{r }
df <- data.frame(
  x = c(5, 1, 4, NA, NA, NA, 2, 1),
  y = c(5, NA, 3, 2, 5, NA, 2, 1),
  z = c(1, 4, 2, 5, 2, NA, 5, 4),
  t = c(4, 2, NA, 1, 5, 2, 3, 3)
)

df

```

Using collaborative filtering, User 2's predicted preference for Movie 6 is approximately **2**.

# Problem 2

```{r}
correlations <- sapply(1:nrow(df), function(i) {
  cor(df[1, ], df[i, ])
})

correlations
```

The movie with the highest Pearson correlation with Movie 1 is:

```{r}
which.max(correlations[-1]) + 1
```

Therefore, the most similar movie to Movie 1 is **Movie 7**.

# Problem 3A

```{r}
countries <- read.csv(
  "http://manctsui.github.io/datasets/countries.csv"
)

for (column in c("Agriculture", "Industry")) {
  
  regional_mean <- ave(
    countries[[column]],
    countries$Region,
    FUN = function(x) mean(x, na.rm = TRUE)
  )
  
  countries[[column]][is.na(countries[[column]])] <-
    regional_mean[is.na(countries[[column]])]
}

countries
```

The missing values in `Agriculture` and `Industry` were replaced with the mean for the corresponding region.

# Problem 3B

```{r}
countries_original <- read.csv(
  "http://manctsui.github.io/datasets/countries.csv"
)

agriculture_na <- table(
  countries_original$Region[
    is.na(countries_original$Agriculture)
  ]
)

industry_na <- table(
  countries_original$Region[
    is.na(countries_original$Industry)
  ]
)

agriculture_na
industry_na
```

The tables above give the number of missing values filled from each region for Agriculture and Industry.

# Problem 4

```{r}
df2 <- data.frame(
  temp = c(79, 74, 73, 76, 82, 72, 71, 79, 83, 70, 73, 75, 77),
  chirps = c(44, 40, 38, 43, 48, 37, 34, 45, 51, 32, 35, 42, 43)
)

r <- cor(df2$temp, df2$chirps)
mean_temp <- mean(df2$temp)
mean_chirps <- mean(df2$chirps)
sd_temp <- sd(df2$temp)
sd_chirps <- sd(df2$chirps)

b1 <- r * (sd_temp / sd_chirps)
b0 <- mean_temp - b1 * mean_chirps

r
mean_temp
mean_chirps
sd_temp
sd_chirps
b1
b0
```

For simple linear regression,

$$
b_1 = r\frac{s_y}{s_x}
$$

and

$$
b_0 = \bar{y} - b_1\bar{x}.
$$

Using the data provided, this gives approximately:

$$
\text{temp} = 0.793(\text{chirps}) + 42.565.
$$

This is the same regression equation obtained from:

```{r}
lm(temp ~ chirps, data = df2)
```

The equation stated in the assignment, `temp = 0.714(chirps) + 46.474`, does not match the data provided.
